Memory-Efficient Mixed-Precision Implementations\\for Robust Explicit Model Predictive Control

### 2019-07-12 emsoft-2019/camera_ready/response.tex

\subsection{Responses}
Here, we provide answer to the main inquiries posed by the reviewers. These points will be added later into the main body of the final version of our paper.\\
\textbf{Significance of memory savings reported in the paper} (Reviewer 1)\\
Explicit MPC is a class of controllers which is mainly appropriate for implementation over processors with limited computational power as it omits the need for massive online optimizations. This speed up in execution comes however with relatively high memory demand due to high number of PWA mappings that must be saved over the memory. Typical memory of these processors is in the order of tens of kilobytes (e.g. $32$ KB) and hence $21$ KB is considered as a significant saving.\\

@@

\textbf{Do memory savings in our approach come with a run-time cost?} (Reviewer 1)\\
No. It should be noticed that savings are not affecting the run-time as synthesizing of the controller and the corresponding mixed precision computations are all performed offline. The output of our analysis is an implementation in mixed precision arithmetic that checks which polyhedral region the current state belongs to and computes the corresponding affine mapping. This computation is fast. See below for further discussion.\\

@@

\textbf{Complexity of the proposed approach} (Reviewer 2)\\
In \cite{Alessio2009} it is shown that the upper bound over the number of regions $P$ is $2^q$, where $q$ denotes the total number of constraints in \autoref{eq:RMPC_prob}. In theory one cannot find an upper bound for the maximum number of hyperplanes enclosing a region. To compute the incorrect region selection error, assuming that each partition at most has $\bar n_f$ faces, in the worst case $i\times\bar n_f\times 2^q$ affine computations would be enough, where $i$ denotes the number of times that RMPC design is repeated for updated $\Delta$s. In our experiments, computation of incorrect region selection takes only a few minutes for our largest tested benchmarks. %\underline{\textbf{Rocco: What about complexity of Daisy's computations in terms of number of regions $P$ and $n$?}} 		
\\
The complexity of precision tuning is linear with respect to the number of regions and hyperplanes. So, $P+\bar n_f\times P$ is an upper bound for the number of problems we need to encode in Daisy.
In each encoding, the uniform-precision detection takes linear number of steps before convergence. It starts from 1-bit and increases the precision one by one until the first valid configuration. The complexity of mixed-precision tuning, with delta-debugging, is $n\times log(n)$ in average, and bounded by $n^2$ in the worst case, with n number of variables.\\

@@

\textbf{Releasing the developed package} (Reviewers 3 $\&$ 4)\\
We definitely plan to distribute our package once the double blinded review phase is passed.\\ 
\textbf{Reason for the special format of solution to \autoref{eq:RMPC_prob} as polyhedral decomposition} (Reviewer 3)\\
We use the results in \cite{delaPea:2005}; since this is not among this paper's contributions, we do not go beyond giving a general outline of these results. In particular, we omitted the proof from \cite{delaPea:2005}.

@@

\end{itemize}
Evaluating the affine map is straightforward and requires less computational effort compared to the region selection step. Direct implementation of region selection (\autoref{lst:caseof}) is simple, but inefficient. More efficiency can be achieved by leveraging binary search tree structures (e.g. \cite{Mnnigmann:2011}). In general, explicit MPC is well-known for its high-speed implementation over embedded systems with low computational power. \cite{Johansen:2007} shows that explicit MPCs with hundreds of regions can be implemented on application specific integrated circuit (ASIC) with about $20000$ gates, leading to computation times in the order of $1\mu s$. In \cite{Bemporad:2006}, it is shown that for typical problems evaluating the explicit MPC takes significantly shorter time in comparison to solving on-line QP. \\
\textbf{Termination of the algorithm for robust EMPC design (Reviewer 4)}\\
From a theoretical point of view, the numerical error does not necessarily converge as each new designed controller may differ from the previous ones. Therefore, it is not easy to give an upper bound over the number of times that the robust MPC design needs to be repeated. However, one can use results on continuity of $argmin$ function in 

@@

%\quad &u_i=\mu x_i+v_i, \quad\forall i\in\{0,1,\cdots,N-1\}\nonumber\\
&u_i\in\mathcal{U},x_i\in\mathcal{X},\quad \forall w_i\in\mathcal{W}_{\Delta},\,\,\forall i\in\{0,1,\cdots,N\},
\end{align*}
 assuming its unique solution to every $\Delta$ and compute the Lipschitz constant for it over variations of $\Delta$. The computed Lipschitz constant will be only dependent on the dynamics of the system depicted in \autoref{eq:DSS} and hence can be used to evaluate the maximum error over incorrect region selection. This Lipschitz constant can be quite large, leading into a very conservative bound on maximum number of iterations before convergence. However, from an experimental point of view, using a reasonable $\varepsilon_{SAFE}$, our observations show that convergence always happened after only a few iterations.

### 2019-07-12 emsoft-2019/camera_ready/Cont_Syn.tex
camera_ready folder is added

\begin{figure}
\begin{lstlisting}[language=Python,numbers=left,numbersep=3pt,frame=lines,keepspaces=true,mathescape=true,basicstyle=\small\ttfamily,escapeinside={<@}{@>}]
 model = input() 
 $\Delta$ = input()
 $\varepsilon_{Q}$ = input()
 assert ($\Delta$ >= 0 and $\varepsilon_{Q}$ > 0)
 $\varepsilon$ = $\varepsilon_{A/D}$ + $\varepsilon_{Q}$<@\Mahmoud{//the width of the uncertain area}@>

>>>>

\begin{figure}
\begin{lstlisting}[language=Python,numbers=left,numbersep=3pt,frame=lines,keepspaces=true,mathescape=true,basicstyle=\small\ttfamily,escapeinside={<@}{@>}]
 model = input() 
 $\Delta$ = input()
 $\varepsilon_{Q}$ = input()
 assert ($\Delta$ >= 0 and $\varepsilon_{Q}$ > 0)
 $\varepsilon$ = $\varepsilon_{A/D}$ + $\varepsilon_{Q}$<@\Mahmoud{//width of the tube}@>

### 2019-07-12 emsoft-2019/camera_ready/finite_precision.tex
camera_ready folder is added

% introduces roundoff errors, these can be tracked with a dataflow analysis
Finite precision unavoidably introduces roundoff errors, which can accumulate
during the course of a computation.
A sound static roundoff error analysis computes a guaranteed upper bound on the
error of the result by tracking worst-case errors at every operation. Given a
word length for each value, several tools\Eva{, including the tool Daisy which we use,} automatically determine the number of
integer bits needed and compute roundoff errors using a dataflow analysis~\cite{Daisy,Gappa}.
They track real-valued ranges at every intermediate operation. These ranges
determine the fixed-point formats and thus also the individual worst-case roundoff errors,
which the analyses track separately from the ranges. To ensure guaranteed
error bounds, both the ranges and errors are computed using interval~\cite{intervals} or
affine arithmetic~\cite{affineArithmetic} which compute sound enclosures. 
Affine arithmetic tracks linear correlations between variables, so for linear expressions
the computed ranges are as tight as possible (nonlinear arithmetic
leads to over-approximations).
% A standard technique to increase accuracy of the computed errors is interval
% subdivision, which divides the input domain into equally-sized subdomains and
% runs roundoff error analysis separately for each subdomain. The
% over-approximations on each subdomain tend to be smaller, which increases the
% overall tightness of range and error bounds.

% mixed-precision tuning
To reduce memory usage and increase efficiency, we want to choose as short word
lengths as possible, \Mahmoud{as this way memory footprint is minimized and run-time is not influenced.} Mixed-precision tuning tools automatically determine possibly different
word lengths for different values~\cite{DaisyTuning,Lee2006}. Compared to uniform precision (or word length),
mixed-precision often leads to improved resource usage, but due to the complexity of
fixed-point arithmetic and the error analysis, is challenging to do manually.  
Automated mixed-precision tuning tools\Eva{, including Daisy,} usually perform a search: they repeatedly
select a candidate mixed-precision assignment (i.e. different word lengths for
different values) and check whether the assignment satisfies a given error
bound. The candidate assignments are chosen based on a heuristic which guides
the search towards promising candidates, e.g. using delta-debugging~\cite{DaisyTuning} or
simulated annealing~\cite{Lee2006}.

>>>>

% introduces roundoff errors, these can be tracked with a dataflow analysis
Finite precision unavoidably introduces roundoff errors, which can accumulate
during the course of a computation.
A sound static roundoff error analysis computes a guaranteed upper bound on the
error of the result by tracking worst-case errors at every operation. Given a
word length for each value, several tools\Mahmoud{, including the tool Daisy which we use,} automatically determine the number of
integer bits needed and compute roundoff errors using a dataflow analysis~\cite{Daisy,Gappa}.
They track real-valued ranges at every intermediate operation. These ranges
determine the fixed-point formats and thus also the individual worst-case roundoff errors,
which the analyses track separately from the ranges. To ensure guaranteed
error bounds, both the ranges and errors are computed using interval~\cite{intervals} or
affine arithmetic~\cite{affineArithmetic} which compute sound enclosures. 
Affine arithmetic tracks linear correlations between variables, so for linear expressions
the computed ranges are as tight as possible (nonlinear arithmetic
leads to over-approximations).
% A standard technique to increase accuracy of the computed errors is interval
% subdivision, which divides the input domain into equally-sized subdomains and
% runs roundoff error analysis separately for each subdomain. The
% over-approximations on each subdomain tend to be smaller, which increases the
% overall tightness of range and error bounds.

% mixed-precision tuning
To reduce memory usage and increase efficiency, we want to choose as short word
lengths as possible, \Mahmoud{as this way memory footprint is minimized and run-time is not influenced.} Mixed-precision tuning tools automatically determine possibly different
word lengths for different values~\cite{DaisyTuning,Lee2006}. Compared to uniform precision (or word length),
mixed-precision often leads to improved resource usage, but due to the complexity of
fixed-point arithmetic and the error analysis, is challenging to do manually.  
Automated mixed-precision tuning tools\Mahmoud{, including Daisy,} usually perform a search: they repeatedly
select a candidate mixed-precision assignment (i.e. different word lengths for
different values) and check whether the assignment satisfies a given error
bound. The candidate assignments are chosen based on a heuristic which guides
the search towards promising candidates, e.g. using delta-debugging~\cite{DaisyTuning} or
simulated annealing~\cite{Lee2006}.
